A re-examination of the role of friction in the original Social Force Model

### 2019-06-22 paper.tex
saque el texto rojo

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics were developed. The class of force-based model was proposed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models assimilate the interaction between 
pedestrians and their environment using different kinds of forces inspired on physical systems. For 
example, Okazaki (1979) modeled the movement of pedestrians as 
a magnetized object immersed a magnetic in field \cite{Okazaki}.}\\

By the late '90s and the beginning of the century, Helbing 
\textit{et al.} proposed a model that included socio-psychological and physical 
forces to simulate crowd dynamics. According to Helbing \textit{et al.}, both 
the environment and the individuals' own desire affect the pedestrians motion in 
a similar way as forces do with respect to the momentum of particles 
\cite{Helbing1,Helbing4}. This ``Social Force Model'' (SFM) related the 
socio-psychological phenomenon of crowds behavior to the ``microscopic'' 
formalism of moving particles. The model succeeded at this instance to 
reproduce the slow down in the efficiency in the evacuation process as 
pedestrians try harder to escape from a dangerous situation (\textit{i.e.} 
``faster is slower'' effect) \cite{Helbing1,Dorso2}.
\textcolor{red}{The faster-is-slower effect was well explained by the presence of human blocking clusters 
that obstruct the exit during an emergency evacuation \cite{Dorso1}.}  \\

@@

Researchers turned back to examine the available data on the velocity and flux 
behavior for different density environments \cite{Boltes,helbing3,seyfried1,seyfried}. 
Ref.~\cite{helbing3} \textcolor{red}{summarizes} empirical data from 
literature, and their own data set, acquired from videos of the Muslim 
pilgrimage in Mina-Makkah (2006). They showed from the empirical fundamental 
diagram (flux $J$ versus density $\rho$) that highly dense crowds (seemingly up 
to $10\,$p~m$^{-2}$) do not drive the pedestrians velocity to zero, although the 
reasons for this remain rather obscure.   \\

\textcolor{red}{More recent findings on the fundamental diagram for an extremely dense
 situation show that, sometimes, the flux may increase with the density (see 
Ref.~\cite{lohner1}). This seems to contradict the results of 
Ref.~\cite{helbing3}. But both results appear not to be completely  comparable 
since data was acquired at different points, and thus,
at different stages of the ritual. Ref.~\cite{lohner1}, indeed, states that the fundamental 
diagram may strongly depend on the type of motion and the nature of the event.}\\

\textcolor{red}{Our work focuses on data from Ref.~\cite{helbing3}. 
Further measurements can be found in 
Refs.~\cite{Boltes,lohner1,Dridi1,Dridi2,Baqui1}.
The PedFlow model developed by L\"ohner \textit{et al}. Ref.~\cite{lohner2}  
is an alternative to the Helbing model. Togashi \textit{et al}. improved the 
calibration of this model using a Kalman filter ensemble \cite{Togashi1}. }\\

@@

Our research was carried out in the context of the ``Social Force Model'' (SFM) 
proposed by Helbing \textcolor{red} {\textit{et al.}} \cite{Helbing1}. This 
model states that human motion is caused by the desire of people to reach a 
certain destination, as well as other environmental factors. The pedestrians 
behavioral pattern in a crowded environment can be modeled by three kind of 
forces: the ``desire force'', the ``social force'' and the ``granular force''.\\

@@

It is well known that the \textcolor{red}{original} version of the Social Force 
Model  is incapable of reproducing the 
fundamental diagram at high densities (say above 
\textcolor{red}{5~p~m$^{-2}$)} \textcolor{red}{\cite{parisi2}} . Different 
approaches were proposed to fix this drawback: increasing the net-time 
headway~\cite{johansson}, canceling the desired velocity~\cite{parisi2} or even 
inducing the jamming state by an attraction target~\cite{kwak}. These approaches 
seems suitable when individuals are not so anxious to reach a certain 
destination. However, when individuals are escaping or rushing due to a 
stressful situation, pedestrians would neither consider reducing their desired 
force nor being captured by an attraction. \\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./corridor.eps}
\caption{\label{corridor} Schematic diagram for individuals in the corridor. 
The circles represent pedestrians moving from left to right. $w$ represents the 
corridor width, $L$ represents the length. The rectangular boxes are upper and 
lower blocks that represent the walls of the corridor. The dashed circle in the 
middle corresponds to the \textcolor{red}{fundamental diagram} measurement region.}
\end{figure}

@@

\section{\label{simulations}\textcolor{red}{Setting and parameters}}

The simulations process were performed on a straight corridor of length $L=$ 
28~m (with periodic boundary conditions) and variable width $w$. We explored 
widths ranging from $w=2$~m to $w=40$~m. The corridor had two side walls, placed 
at $y=0$ and $y=w$, respectively. The length of each wall was $L$. The 
pedestrians were modeled as soft spheres of radius $R_i=0.23$~m. This size was 
fixed according to Ref.~\cite{metric_handbook}. Initially, the individuals were 
randomly distributed along the corridor with a fixed global density 
(\textcolor{red}{p~m$^{-2}$}) and with random initial velocities, resembling a 
Gaussian distribution with null mean value. We explored global density values in 
the range 1~\textcolor{red}{p~m$^{-2}$}~$<\rho<$~9~\textcolor{red}{p~m$^{-2}$}. 
We did not explore extreme densities (say $\rho>$9~\textcolor{red}{p~m$^{-2}$}) 
because we excluded injuring situations. The number of pedestrians in the 
simulation was given by the global density and the corridor dimensions chosen in 
each case. \\

@@

The desired velocity for each pedestrian $i$ was 
$\vec{v}_d^{~(i)}=1$~m~s$^{-1}$~$\hat{e}_d^{~(i)}$, where the target $\hat{e}_d^{~(i)}$ 
was set as $\hat{e}_d^{~(i)}=(L,y_i)\left \| (L,y_i) \right \|^{-1}$, being $L$ 
the \textit{x}-location of the end at the corridor and $y_i$ the 
\textit{y}-location corresponding to the \textit{ith} pedestrian 
(\textcolor{red}{see Fig.~\ref{corridor}}). This allowed the pedestrians to 
move from left to right in an unidirectional flow. Pedestrians that surpassed 
$x=L$ were re-injected at $x=0$, preserving their current velocity and 
\textit{y}-location (\textit{i.e.} periodic boundary conditions). This mechanism 
was carried out in order to keep
the crowd size unchanged.\\

@@

\subsection{\label{software}\textcolor{red}{Simulation software}}

@@

\section{\label{Hypotheses} \textcolor{red}{Hypotheses}}

@@

\textcolor{red}{Our starting point will be the re-examination of the (basic) SFM. 
We propose a reduce-in-units equation of motion of the SFM (see Section~\ref{appendix_1} 
for details). From this point of view, the parameter $A$ standing for the intensity of the
 social force $\mathbf{f}_s$ is replaced by the reduced-in-units parameter $\mathcal{A}$.
 The friction intensity $\kappa$ of the granular force $\mathbf{f}_g$ is replaced by the reduced-in-units
 parameter $\mathcal{K}$. No other parameters are required in the reduced-in-units model
 since $m$, $v_d$, $\tau$ and $B$ are all included in $\mathcal{A}$ and $\mathcal{K}$.
 The desired force $\mathbf{f}_d$, indeed, will only depend on the target direction. In Section~\ref{appendix_1} we derivate the 
reduce-in-units equation of motion and the meaning of the control parameters ($\mathcal{A}$ and $\mathcal{K}$). }\\

@@

The former ``re-calibrations" accomplish the socio-psychological
response of the crowd to ``gain more space" (by either enhancing local
interactions or performing a delayed reaction). We are aware that
crowds may respond differently in many situations (see Ref.~\cite{drury1}). Our working
hypothesis, however, does not focus on the crowd socio-psychological
response, but on the physical contact among pedestrians (and the
walls). The socio-psychological attitude of the pedestrians will be
assumed to remain fixed along the simulations (with the 
\textcolor{red}{desired speed} limited to $v_d=1\,$ m~s$^{-1}$). \\

@@

\begin{equation}
\textcolor{red}{v(\rho_{max}) < \frac{5v_d}{\rho_{max}} \leq \frac{5}{9} v_d}\\
\end{equation}

@@

We may summarize our first results as follows. We were able to validate the 
\textcolor{red}{original} SFM for narrow corridors through the fundamental 
diagram. However, the SFM (in its current version) disagrees with experimental 
data as the corridors widen. We will focus in the next Section on the velocity 
profile in order to investigate this discrepancy.

@@

As we mentioned in Section \ref{fundamental_diagram}, when the density is low, pedestrians achieve the desired velocity ($v=v_d=1$~m~s$^{-1}$). Since the results of the previous Section only hold for the area located in the middle of the corridor (see \textcolor{red}{dashed circle in} Fig.~\ref{corridor}), we want to shed some light and understand what is happening across the entire corridor.\\

We first noticed that low-density situations ($\rho<5$) lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities($\rho>5$), the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution. \textcolor{red}{The velocity profile is in agreement with the empirical data in Ref.~\cite{zhang1}.}  \\

@@

\textcolor{red}{We envisage $\mathcal{A}$ as a parameter related exclusively to the 
repulsion between pedestrians and $\mathcal{K}$  as parameter related to the friction-repulsion.
While the former is valid in every situation, the latter only takes part when the
pedestrians are in contact.}\\

@@

As already mentioned, the results shown so far indicate that friction may be the 
key magnitude for fitting the fundamental diagram into the experimental data. We 
want to make clear that fitting the experimental data means mimicking 
(qualitatively) the congested regime reported by different authors (including 
the Jamaraat study in Ref.~\cite{helbing3}) for corridors as width as 22~m. The 
\textcolor{red}{original} version of the Social Force Model proposes the same 
friction coefficient for the pedestrian-pedestrian interaction and the 
pedestrian-wall interaction. The proposed value was $\kappa = 2.4\times10^{5}$. 
This value is widely used in many studies.  \\

@@

Cluster formation is a very important process in pedestrian dynamics. Moreover, is the key process that explains the clogging phenomena in bottleneck evacuations. We analyzed the clustering formation according to the granular cluster definition given in Section~\ref{granular-cluster}. Fig.~\ref{cluster_distribution} shows the histograms of the cluster size distribution for three different densities, from top to bottom: $\rho=4.5$, $\rho=5$ and $\rho=5.5$. These three
densities are representative for the crossover between
a non-clusterized regime and a unique “giant” cluster
regime. We studied two situations for each density: the
\textcolor{red}{original} SFM on left hand side plots, and the enhanced
friction situation on the right hand side plots (see
caption for details).\\

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many ``caged" in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).
\textcolor{red}{In other words, the bimodal distribution occurs because,
after the giant component is formed, many small clusters remain ``caged" 
inside the giant component. These small clusters (or single individuals)
do not touch permanently any pedestrian belonging to the giant component.} \\

@@

We further propose modeling the pedestrian-wall friction 
interaction with a different coefficient than the pedestrian-pedestrian friction 
interaction. This does neither mean a ``re-calibration'' (for 
high dense crowds) nor a departure from the \textcolor{red}{original} SFM model. 
We actually find no reason for changing other parameters, regardless of any 
specific situation. We also want to stress the fact that studying the friction 
coefficients may be a critical factor to properly reproduce the dynamics of a 
massive evacuation under high levels of anxiety. 
\\

@@

  \bibitem{Hirai}
\textcolor{red}{Hirai, K., and Tarui, K., 1975. A simulation of the behavior of 
a crowd in panic, Proceedings of the 1975 International
Conference on Cybernetics Society, pp. 409-411}

 \bibitem[Okazaki (1979)]{Okazaki}
\textcolor{red}{Okazaki, S., 1979. A study of pedestrian movement in 
architectural space, Part 
1: pedestrian movement by the application of magnetic model. Trans. A.I.J. 283, 
111–119.}

@@

\bibitem{koster1}
\textcolor{red}{K\"oster}, G., Treml, F. and G\"odel, M., 2013. "Avoiding 
numerical pitfalls in social force models." Physical Review E 87 (6), 063305.
{\path{https://doi.org/10.1103/PhysRevE.87.063305}}

@@

\bibitem{dietrich1}
Dietrich, F., \textcolor{red}{K\"oster}, G., Seitz, M. and 
von Sivers, I., 2014. "Bridging the gap: From cellular automata to 
differential equation models for pedestrian dynamics." Journal of Computational 
Science 5 (5), 841-846.{\path{https://doi.org/10.1016/j.jocs.2014.06.005}}

@@

\bibitem{Boltes}
\textcolor{red}{Boltes, M., Zhang, J., Tordeux, A., Schadschneider A., Seyfried 
A., 
2018. Empirical Results of Pedestrian and Evacuation Dynamics. In: Meyers R. 
(eds) Encyclopedia of Complexity and Systems Science. Springer, Berlin, 
Heidelberg}

@@

\bibitem{lohner1}
\textcolor{red}{L\"ohner}, R., Muhamad, B., Dambalmath, P., and Haug, E., 2018. 
"Fundamental diagrams for specific very high density crowds." Collective 
Dynamics 2, 1-15.
{\path{http://dx.doi.org/10.17815/CD.2017.13}}

\bibitem{Dridi1}
\textcolor{red}{Dridi, M. (2015) Pedestrian Flow Simulation Validation and
Verification Techniques. Current Urban Studies, 3, 119-134.
doi: 10.4236/cus.2015.32011.}

\bibitem{Dridi2}
\textcolor{red}{Dridi, M. (2015) Simulation of High Density Pedestrian Flow:
A Microscopic Model. Open Journal of Modelling and Simulation,
3, 81-95. doi: 10.4236/ojmsi.2015.33009.}

\bibitem{Baqui1}
\textcolor{red}{Baqui, M., and L\"ohner, R. "Real-time crowd safety
and comfort management from CCTV images." Real-Time Image and Video Processing 2017.
Vol. 10223. International Society for Optics and Photonics, 2017.}

\bibitem{lohner2}
\textcolor{red}{L\"ohner, R. "On the modeling of pedestrian motion." 
Applied Mathematical Modelling 34.2 (2010): 366-382.}

\bibitem{Togashi1}
\textcolor{red}{Togashi, F., Misaka, T., L\"ohner, R., and Obayashi, S. (2018). 
Using ensemble Kalman filter to determine parameters for computational crowd dynamics simulations. 
Engineering Computations, 35(7), 2612-2628.}

@@

\bibitem{zhang1}
\textcolor{red}{Zhang, X. L., Weng, W. G., Yuan, H. Y., and Chen, J. G. (2013). 
Empirical study of a unidirectional dense crowd during a real mass event. 
Physica A: Statistical Mechanics and its Applications, 392(12), 2781-2791.}

>>>>

In the last decades, many microscopic models 
for crowd dynamics were developed. The class of force-based model was proposed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models assimilate the interaction between 
pedestrians and their environment using different kinds of forces inspired on physical systems. For 
example, Okazaki (1979) modeled the movement of pedestrians as 
a magnetized object immersed a magnetic in field \cite{Okazaki}.\\

By the late '90s and the beginning of the century, Helbing 
\textit{et al.} proposed a model that included socio-psychological and physical 
forces to simulate crowd dynamics. According to Helbing \textit{et al.}, both 
the environment and the individuals' own desire affect the pedestrians motion in 
a similar way as forces do with respect to the momentum of particles 
\cite{Helbing1,Helbing4}. This ``Social Force Model'' (SFM) related the 
socio-psychological phenomenon of crowds behavior to the ``microscopic'' 
formalism of moving particles. The model succeeded at this instance to 
reproduce the slow down in the efficiency in the evacuation process as 
pedestrians try harder to escape from a dangerous situation (\textit{i.e.} 
``faster is slower'' effect) \cite{Helbing1,Dorso2}.
The faster-is-slower effect was well explained by the presence of human blocking clusters 
that obstruct the exit during an emergency evacuation \cite{Dorso1}.  \\

@@

Researchers turned back to examine the available data on the velocity and flux 
behavior for different density environments \cite{Boltes,helbing3,seyfried1,seyfried}. 
Ref.~\cite{helbing3} summarizes empirical data from 
literature, and their own data set, acquired from videos of the Muslim 
pilgrimage in Mina-Makkah (2006). They showed from the empirical fundamental 
diagram (flux $J$ versus density $\rho$) that highly dense crowds (seemingly up 
to $10\,$p~m$^{-2}$) do not drive the pedestrians velocity to zero, although the 
reasons for this remain rather obscure.   \\

More recent findings on the fundamental diagram for an extremely dense
 situation show that, sometimes, the flux may increase with the density (see 
Ref.~\cite{lohner1}). This seems to contradict the results of 
Ref.~\cite{helbing3}. But both results appear not to be completely  comparable 
since data was acquired at different points, and thus,
at different stages of the ritual. Ref.~\cite{lohner1}, indeed, states that the fundamental 
diagram may strongly depend on the type of motion and the nature of the event.\\

Our work focuses on data from Ref.~\cite{helbing3}. 
Further measurements can be found in 
Refs.~\cite{Boltes,lohner1,Dridi1,Dridi2,Baqui1}.
The PedFlow model developed by L\"ohner \textit{et al}. Ref.~\cite{lohner2}  
is an alternative to the Helbing model. Togashi \textit{et al}. improved the 
calibration of this model using a Kalman filter ensemble \cite{Togashi1}. \\

@@

Our research was carried out in the context of the ``Social Force Model'' (SFM) 
proposed by Helbing \textit{et al.} \cite{Helbing1}. This 
model states that human motion is caused by the desire of people to reach a 
certain destination, as well as other environmental factors. The pedestrians 
behavioral pattern in a crowded environment can be modeled by three kind of 
forces: the ``desire force'', the ``social force'' and the ``granular force''.\\

@@

It is well known that the original version of the Social Force 
Model  is incapable of reproducing the 
fundamental diagram at high densities (say above 
5~p~m$^{-2}$) \cite{parisi2} . Different 
approaches were proposed to fix this drawback: increasing the net-time 
headway~\cite{johansson}, canceling the desired velocity~\cite{parisi2} or even 
inducing the jamming state by an attraction target~\cite{kwak}. These approaches 
seems suitable when individuals are not so anxious to reach a certain 
destination. However, when individuals are escaping or rushing due to a 
stressful situation, pedestrians would neither consider reducing their desired 
force nor being captured by an attraction. \\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./corridor.eps}
\caption{\label{corridor} Schematic diagram for individuals in the corridor. 
The circles represent pedestrians moving from left to right. $w$ represents the 
corridor width, $L$ represents the length. The rectangular boxes are upper and 
lower blocks that represent the walls of the corridor. The dashed circle in the 
middle corresponds to the fundamental diagram measurement region.}
\end{figure}

@@

\section{\label{simulations}Setting and parameters}

The simulations process were performed on a straight corridor of length $L=$ 
28~m (with periodic boundary conditions) and variable width $w$. We explored 
widths ranging from $w=2$~m to $w=40$~m. The corridor had two side walls, placed 
at $y=0$ and $y=w$, respectively. The length of each wall was $L$. The 
pedestrians were modeled as soft spheres of radius $R_i=0.23$~m. This size was 
fixed according to Ref.~\cite{metric_handbook}. Initially, the individuals were 
randomly distributed along the corridor with a fixed global density 
(p~m$^{-2}$) and with random initial velocities, resembling a 
Gaussian distribution with null mean value. We explored global density values in 
the range 1~p~m$^{-2}$~$<\rho<$~9~p~m$^{-2}$. 
We did not explore extreme densities (say $\rho>$9~p~m$^{-2}$) 
because we excluded injuring situations. The number of pedestrians in the 
simulation was given by the global density and the corridor dimensions chosen in 
each case. \\

@@

The desired velocity for each pedestrian $i$ was 
$\vec{v}_d^{~(i)}=1$~m~s$^{-1}$~$\hat{e}_d^{~(i)}$, where the target $\hat{e}_d^{~(i)}$ 
was set as $\hat{e}_d^{~(i)}=(L,y_i)\left \| (L,y_i) \right \|^{-1}$, being $L$ 
the \textit{x}-location of the end at the corridor and $y_i$ the 
\textit{y}-location corresponding to the \textit{ith} pedestrian 
(see Fig.~\ref{corridor}). This allowed the pedestrians to 
move from left to right in an unidirectional flow. Pedestrians that surpassed 
$x=L$ were re-injected at $x=0$, preserving their current velocity and 
\textit{y}-location (\textit{i.e.} periodic boundary conditions). This mechanism 
was carried out in order to keep
the crowd size unchanged.\\

@@

\subsection{\label{software} Simulation software}

@@

\section{\label{Hypotheses} Hypotheses}

@@

Our starting point will be the re-examination of the (basic) SFM. 
We propose a reduce-in-units equation of motion of the SFM (see Section~\ref{appendix_1} 
for details). From this point of view, the parameter $A$ standing for the intensity of the
 social force $\mathbf{f}_s$ is replaced by the reduced-in-units parameter $\mathcal{A}$.
 The friction intensity $\kappa$ of the granular force $\mathbf{f}_g$ is replaced by the reduced-in-units
 parameter $\mathcal{K}$. No other parameters are required in the reduced-in-units model
 since $m$, $v_d$, $\tau$ and $B$ are all included in $\mathcal{A}$ and $\mathcal{K}$.
 The desired force $\mathbf{f}_d$, indeed, will only depend on the target direction. In Section~\ref{appendix_1} we derivate the 
reduce-in-units equation of motion and the meaning of the control parameters ($\mathcal{A}$ and $\mathcal{K}$). \\

@@

The former ``re-calibrations" accomplish the socio-psychological
response of the crowd to ``gain more space" (by either enhancing local
interactions or performing a delayed reaction). We are aware that
crowds may respond differently in many situations (see Ref.~\cite{drury1}). Our working
hypothesis, however, does not focus on the crowd socio-psychological
response, but on the physical contact among pedestrians (and the
walls). The socio-psychological attitude of the pedestrians will be
assumed to remain fixed along the simulations (with the 
desired speed limited to $v_d=1\,$ m~s$^{-1}$). \\

@@

\begin{equation}
v(\rho_{max}) < \frac{5v_d}{\rho_{max}} \leq \frac{5}{9} v_d\\
\end{equation}

@@

We may summarize our first results as follows. We were able to validate the 
original SFM for narrow corridors through the fundamental 
diagram. However, the SFM (in its current version) disagrees with experimental 
data as the corridors widen. We will focus in the next Section on the velocity 
profile in order to investigate this discrepancy.

@@

As we mentioned in Section \ref{fundamental_diagram}, when the density is low, pedestrians achieve the desired velocity ($v=v_d=1$~m~s$^{-1}$). Since the results of the previous Section only hold for the area located in the middle of the corridor (see dashed circle in Fig.~\ref{corridor}), we want to shed some light and understand what is happening across the entire corridor.\\

We first noticed that low-density situations ($\rho<5$) lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities($\rho>5$), the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution. The velocity profile is in agreement with the empirical data in Ref.~\cite{zhang1}.  \\

@@

We envisage $\mathcal{A}$ as a parameter related exclusively to the 
repulsion between pedestrians and $\mathcal{K}$  as parameter related to the friction-repulsion.
While the former is valid in every situation, the latter only takes part when the
pedestrians are in contact.\\

@@

As already mentioned, the results shown so far indicate that friction may be the 
key magnitude for fitting the fundamental diagram into the experimental data. We 
want to make clear that fitting the experimental data means mimicking 
(qualitatively) the congested regime reported by different authors (including 
the Jamaraat study in Ref.~\cite{helbing3}) for corridors as width as 22~m. The 
original version of the Social Force Model proposes the same 
friction coefficient for the pedestrian-pedestrian interaction and the 
pedestrian-wall interaction. The proposed value was $\kappa = 2.4\times10^{5}$. 
This value is widely used in many studies.  \\

@@

Cluster formation is a very important process in pedestrian dynamics. Moreover, is the key process that explains the clogging phenomena in bottleneck evacuations. We analyzed the clustering formation according to the granular cluster definition given in Section~\ref{granular-cluster}. Fig.~\ref{cluster_distribution} shows the histograms of the cluster size distribution for three different densities, from top to bottom: $\rho=4.5$, $\rho=5$ and $\rho=5.5$. These three
densities are representative for the crossover between
a non-clusterized regime and a unique “giant” cluster
regime. We studied two situations for each density: the
original SFM on left hand side plots, and the enhanced
friction situation on the right hand side plots (see
caption for details).\\

We found two unexpected results. Increasing the density produces bigger size clusters until the size distribution suddenly switches to a bimodal distribution (compare \ref{size_distribution_w22_density5} with \ref{size_distribution_w22_density5_5} and \ref{size_distribution_w22_density4.5_kx10} with \ref{size_distribution_w22_density5_kx10}). Once
this phenomenon occurs, any of two possibilities may
appear: the pedestrian belongs to a small cluster (out
of many ``caged" in the crowd) or he (she) belongs to the
giant cluster (with a size comparable to the entire crowd).
In other words, the bimodal distribution occurs because,
after the giant component is formed, many small clusters remain ``caged" 
inside the giant component. These small clusters (or single individuals)
do not touch permanently any pedestrian belonging to the giant component. \\

@@

We further propose modeling the pedestrian-wall friction 
interaction with a different coefficient than the pedestrian-pedestrian friction 
interaction. This does neither mean a ``re-calibration'' (for 
high dense crowds) nor a departure from the original SFM model. 
We actually find no reason for changing other parameters, regardless of any 
specific situation. We also want to stress the fact that studying the friction 
coefficients may be a critical factor to properly reproduce the dynamics of a 
massive evacuation under high levels of anxiety. 
\\

@@

  \bibitem{Hirai}
Hirai, K., and Tarui, K., 1975. A simulation of the behavior of 
a crowd in panic, Proceedings of the 1975 International
Conference on Cybernetics Society, pp. 409-411

 \bibitem[Okazaki (1979)]{Okazaki}
Okazaki, S., 1979. A study of pedestrian movement in 
architectural space, Part 
1: pedestrian movement by the application of magnetic model. Trans. A.I.J. 283, 
111–119.

@@

\bibitem{koster1}
K\"oster, G., Treml, F. and G\"odel, M., 2013. "Avoiding 
numerical pitfalls in social force models." Physical Review E 87 (6), 063305.
{\path{https://doi.org/10.1103/PhysRevE.87.063305}}

@@

\bibitem{dietrich1}
Dietrich, F., K\"oster, G., Seitz, M. and 
von Sivers, I., 2014. "Bridging the gap: From cellular automata to 
differential equation models for pedestrian dynamics." Journal of Computational 
Science 5 (5), 841-846.{\path{https://doi.org/10.1016/j.jocs.2014.06.005}}

@@

\bibitem{Boltes}
Boltes, M., Zhang, J., Tordeux, A., Schadschneider A., Seyfried 
A., 
2018. Empirical Results of Pedestrian and Evacuation Dynamics. In: Meyers R. 
(eds) Encyclopedia of Complexity and Systems Science. Springer, Berlin, 
Heidelberg

@@

\bibitem{lohner1}
L\"ohner, R., Muhamad, B., Dambalmath, P., and Haug, E., 2018. 
"Fundamental diagrams for specific very high density crowds." Collective 
Dynamics 2, 1-15.
{\path{http://dx.doi.org/10.17815/CD.2017.13}}

\bibitem{Dridi1}
Dridi, M. (2015) Pedestrian Flow Simulation Validation and
Verification Techniques. Current Urban Studies, 3, 119-134.
doi: 10.4236/cus.2015.32011.

\bibitem{Dridi2}
Dridi, M. (2015) Simulation of High Density Pedestrian Flow:
A Microscopic Model. Open Journal of Modelling and Simulation,
3, 81-95. doi: 10.4236/ojmsi.2015.33009.

\bibitem{Baqui1}
Baqui, M., and L\"ohner, R. "Real-time crowd safety
and comfort management from CCTV images." Real-Time Image and Video Processing 2017.
Vol. 10223. International Society for Optics and Photonics, 2017.

\bibitem{lohner2}
L\"ohner, R. "On the modeling of pedestrian motion." 
Applied Mathematical Modelling 34.2 (2010): 366-382.

\bibitem{Togashi1}
Togashi, F., Misaka, T., L\"ohner, R., and Obayashi, S. (2018). 
Using ensemble Kalman filter to determine parameters for computational crowd dynamics simulations. 
Engineering Computations, 35(7), 2612-2628.

@@

\bibitem{zhang1}
Zhang, X. L., Weng, W. G., Yuan, H. Y., and Chen, J. G. (2013). 
Empirical study of a unidirectional dense crowd during a real mass event. 
Physica A: Statistical Mechanics and its Applications, 392(12), 2781-2791.
